\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{enumitem}
\setlist[itemize]{itemsep=0.25em}
\setlist[enumerate]{itemsep=0.4em}

\begin{document}

\section*{Programming Problem 1: LIME for Classification}

Use the \textbf{Diabetes dataset}.

\begin{enumerate}
\item Train three classification models (e.g. Logistic Regression, SVM, and MLP). (3 points)

\item Select 5 test samples and generate local explanations using \textbf{LIME}. (5 points)

\item Report the most influential features identified by LIME for each sample. (3 points)

\item Discuss whether the influential features align with the global model behavior. (5 points)

\item In LIME, the importance of perturbed samples is determined using a kernel function that assigns higher weights to points closer to the instance being explained. The kernel is defined as:

\[
\pi_x(z) = \exp\!\left( - \frac{D(x, z)^2}{(0.75\sqrt{d})^2} \right),
\]

where:
\begin{itemize}
    \item x = instance being explained,
    \item z = perturbed sample,
    \item D(x, z) = Euclidean distance between x and z,
    \item d = number of features in the dataset.
\end{itemize}

Analyze how the choice of the kernel width parameter $0.75\sqrt{d}$ affects the distribution of weights. Experiment by replacing 0.75 with 0.25 and 1.0. Discuss your observations on two samples. (4 points)
\end{enumerate}

\section*{Programming Problem 2: SHAP for Classification}

Use the \textbf{Breast Cancer dataset} from sklearn.datasets.load\_breast\_cancer.

\begin{enumerate}
\item Train three classification models (e.g. Logistic Regression, SVM, and MLP). (3 points)

\item Use \textbf{SHAP} to explain predictions for 5 randomly chosen test samples. (5 points)

\item Plot SHAP summary and dependence plots to visualize feature importance. (5 points)
\end{enumerate}

\section*{Programming Problem 3: LIME and SHAP on MNIST}

Use the \textbf{MNIST dataset}.

\begin{enumerate}
\item Train a CNN model on MNIST. (2 points)

\item Use \textbf{LIME} to explain predictions for 5 test images. (5 points)

\item Use \textbf{SHAP (DeepExplainer or GradientExplainer)} to explain the same 5 images. (5 points)

\item Compare and discuss the differences and similarities between LIME and SHAP explanations. (5 point)
\end{enumerate}

\section*{Programming Problem 4: LIME and SHAP with Pretrained Model}

\begin{enumerate}
\item Load a pretrained \textbf{Inception V3 or ResNet50} model and classify 5 CIFAR-10 images. (5 points)

\item Generate explanations for each image using both \textbf{LIME} and \textbf{SHAP}. (5 points)

\item Critically evaluate which method provides more intuitive explanations and justify your reasoning. (5 points)
\end{enumerate}

\end{document}
